### 2019-01-31 paper/paper.tex
paper.tex: spellcheck. References.

% $Id$
\documentclass[twocolumn,prd,nofootinbib]{revtex4}

\newcommand\ForInternalReference[1]{}
%\newcommand\SkipForEarlyCirculation[1]{}
\newcommand\SkipForEarlyCirculation[1]{#1}
%\newcommand\SkipBoxing[1]{}
\newcommand\SkipBoxing[1]{#1}
\newcommand\SkipPP[1]{}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{color}
\usepackage{xspace}
\usepackage{url}
\usepackage{amsmath}
%\usepackage{adjustbox}
\usepackage{float}
\usepackage{multirow}
%
%
\usepackage{times}
%
%
%
\newcommand\optional[1]{}

%
\newcommand\E[1]{\left\langle #1\right\rangle}
\newcommand\qmstate[1]{\left|#1\right \rangle}
\newcommand\qmstateKet[1]{\left\langle#1\right|}
\newcommand\qmstateproduct[2]{\left\langle#1|#2\right\rangle}
\newcommand\qmoperatorelement[3]{\left\langle#1\left|#2\right|#3\right\rangle}
\newcommand\qmoperator[1]{{\bf #1}}
%
\newcommand\Y[1]{{{}_{#1}Y}}

\newcommand\lnL{ \ln {\cal L}}
\newcommand\lnLmarg{ \ln{\cal L}_{\rm marg}{}}
\newcommand\unit[1]{{\rm #1}}

\newcommand\rapidPEOrig{rapid\_pe1}
\newcommand\ILE{ILE}
\newcommand\editremark[1]{{\color{red} #1}}
%
%
%
\usepackage{color}
\definecolor{amber}{rgb}{1.0, 0.75, 0.0}
\definecolor{orange}{rgb}{1.0, 0.5, 0.0}
\definecolor{amaranth}{rgb}{0.9, 0.17, 0.31}
\def\fixme#1{\textcolor{red}{#1}}
\newcommand{\Richard}[1]{ {\color{blue}{#1}}}
\newcommand{\ros}[1]{ {\color{blue}{#1}}}
%

%

%
%
%
%
\graphicspath{{./figures/}}
\newcommand{\mc}{{\cal M}}
\newcommand{\Ms}{M_{\odot}}
\newcommand\LambdaTilde{\widetilde{\Lambda}}
\newcommand\DeltaLambdaTilde{\delta \widetilde{\Lambda}}
%
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\simlt{\lower.5ex\hbox{\ltsima}}
\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\simgt{\lower.5ex\hbox{\gtsima}}

\newcommand\prx{Phys.~Rev.~X}
\def\aj{Astronomical Journal}                 %
\def\apj{Astrophysical Journal}                %
\def\apjl{Astrophysical Journal}             %
\def\pasj{PASJ}
\def\apjs{ApJS}              %
\def\mnras{MNRAS}            %
\def\prd{Phys.~Rev.~D}       %
\def\prl{Phys.~Rev.~Lett}    %
\def\cqg{Class.~Quant.~Grav.~}%
\def\araa{ARA\&A}             %
\def\nat{Nature}              %
\def\aap{A\&A}                %
\def\aapr{A\&A~Rev.~}    %
\def\pasp{PASP}    %
\def\sovast{Soviet Ast.}
%
%

\newcommand{\IMRPD}{\textsc{IMRPhenomD}\xspace}
\newcommand{\IMRPDT}{\textsc{IMRPhenomD\_NRTidal}\xspace}
\newcommand{\IMRP}{\textsc{IMRPhenomPv2}\xspace}
\newcommand{\SEOBP}{\textsc{SEOBNRv3}\xspace}
\newcommand{\SEOBA}{\textsc{SEOBNRv4}\xspace}
\newcommand{\SEOBAROM}{\textsc{SEOBNRv4\_ROM}\xspace}
\newcommand{\NRSur}{NRSur7dq2\xspace}
\newcommand{\TEOB}{SEOBNRv4T\xspace}
\newcommand{\Resum}{TEOBResumS\xspace}
\newcommand\RIFT{RIFT}
\newcommand{\Taylor}{TaylorF2\xspace}
\newcommand\PaperDetection{\underline{LVC-detect}\cite{DiscoveryPaper}}
\newcommand\PaperPE{\underline{LVC-PE}\cite{PEPaper}}
\newcommand\PaperTestGR{\underline{LVC-TestGR}\cite{TestingGRPaper}}
\newcommand\PaperPENRMethods{\underline{PE+NR-Methods}\cite{gwastro-PENR-Methods-Lange}}
\newcommand\PaperAstro{\underline{LVC-Astro}\cite{AstroPaper}}
\newcommand\PaperBurst{\underline{LVC-Burst}\cite{BurstPaper}}
\newcommand\PaperRates{\underline{LVC-Rates}\cite{RatesPaper}}
\newcommand\PaperStochastic{\underline{LVC-Stochastic}}
\newcommand\PaperSEOBNRvthree{\underline{LVC-SEOBNRv3}\cite{SEOBv3Paper}}

\def\RIT{Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, New York 14623, USA}

\begin{document}

\title{Accelerating parameter inference with graphics processing units}
\author{D. Wysocki}
\affiliation{\RIT}
\author{Yao-Lung L. Fang}
%\author{Yao-Lung L. Fang (方耀龍)}
\affiliation{Computational Science Initiative, Brookhaven National Laboratory, Upton, NY 11973, USA}
\affiliation{National Synchrotron Light Source II, Brookhaven National Laboratory, Upton, NY 11973, USA}
\author{R. O'Shaughnessy}
\affiliation{\RIT}
\author{Jake?}
\affiliation{\RIT}
\begin{abstract}
Gravitational wave Bayesian parameter inference involves repeated comparisons of GW data to generic candidate predictions.
Even with algorithmically efficient methods like RIFT or reduced-order quadrature, the time needed to perform these
calculations and overall computational cost can be significant compared to the minutes to hours needed to achieve the
goals of low-latency multimessenger astronomy.  By translating some elements of the RIFT algorithm to operate on
graphics processing units (GPU), we demonstrate substantial performance improvements, enabling dramatically reduced
overall cost and latency. 
\end{abstract}
\maketitle

\section{Introduction}
The Advanced LIGO  \cite{2015CQGra..32g4001T}  and Virgo \cite{gw-detectors-Virgo-original-preferred}  ground-based gravitational wave (GW) detectors have
identified several coalescing compact binaries
\cite{DiscoveryPaper,LIGO-O1-BBH,2017PhRvL.118v1101A,LIGO-GW170814,LIGO-GW170608,LIGO-GW170817-bns}.  
%
The properties of the sources responsible have been inferred via Bayesian comparison of  each source with candidate
gravitational wave signals
\cite{DiscoveryPaper,LIGO-O1-BBH,2017PhRvL.118v1101A,LIGO-GW170814,LIGO-GW170608,LIGO-GW170817-bns}
\editremark{citations: real events, also LI and RIFT} \cite{gwastro-PE-AlternativeArchitectures,gwastro-PENR-RIFT}.  
Many more are expected as observatories reach design sensitivity \cite{2016LRR....19....1A}.  
Both to handle the rapid pace of discovery and particularly to enable coordinated multimessenger followup, GW
observatories should be able to reconstruct the full source parameters of coalescing binaries as fast as possible
\cite{LIGO-whitepaper-2015} \editremark{update}.    
However, particularly when using the best-available waveform models, these calculations can be very costly.  In
practice, detailed inferences even using simplified waveforms take weeks for binary neutron stars, and can take months
for costly time-domain effective-one-body models which incorporate the effects of precession.

% BORROWED FROM ROM: edit
Several strategies have been
developed to reduce the 
%overall computational cost and turnaround time of 
computational cost of parameter estimation \cite{gwastro-pe-Brandon-STF2,gwastro-PE-AlternativeArchitectures,gw-astro-ReducedOrderQuadraturePE-TiglioEtAl2014,2016PhRvD..94d4031S,2017CQGra..34k5006V}.  
% The computational cost of parameter estimation can be reduced 
Approaches that have appeared in the literature include
%by 
generating the approximate solutions more quickly \cite{gwastro-mergers-PE-ReducedOrder-2013,2014PhRvX...4c1006F,2013PhRvD..87l2002S,2013PhRvD..87d4008C,gwastro-mergers-IMRPhenomP,gwastro-SpinTaylorF2-2013}; 
%by 
interpolating some combination
of the waveform or likelihood \cite{gwastro-approx-ROMNR-Blackman2015,2013PhRvD..87d4008C,2013PhRvD..87l2002S,2014CQGra..31s5010P,2014PhRvD..90d4074S,gwastro-PE-AlternativeArchitectures,cole2014likelihood,2012MNRAS.421..169G};  or 
%by 
adopting a sparse representation to reduce the computational cost of data
handling \cite{antil2013two,gwastro-mergers-PE-ReducedOrder-2013,2016PhRvD..94d4031S,gw-astro-ReducedOrderQuadraturePE-TiglioEtAl2014,gwastro-PE-AlternativeArchitectures}.   % 2014PhRvX...4c1006F,2013PhRvD..87l2002S,2013PhRvD..87d4008C
 Some methods, however, achieve rapid turnaround through simplifying approximations.  
%
The RIFT strategy described in \editremark{citations - include ROM} eschews these simplifications, performing
embarrassingly-parallel inferences even for costly models.   However, the method as previously implemented still had a
significant net  computational cost and noticeable startup time, limiting its diverse potential applications.

In this paper we describe a substantial improvement to the RIFT approach, enabled by translating its low-level
algorithms to operate on graphics processing units (GPUs).

This paper is organized as follows.
In Section \ref{Methods}, we review the underlying marginalized likelihood calculations used by RIFT; their updated GPU
implementation; and the operating settings which enable substantially increased performance.
In Section \ref{sec:demo}, we quantify the improved performance of our  GPU-accelerated code, and its agreement with the
original CPU-only version.

\section{Methods}

\subsection{Parameter inference with the RIFT likelihood}
% From RIFT paper
\ILE{}  -- a specific algorithm to ``integrate over extrinsic parameters'' -- provides a straightforward and efficient mechanism to compare any specific candidate gravitational wave source with
real or synthetic data   \cite{gwastro-PE-AlternativeArchitectures,NRPaper,2017PhRvD..96j4041L,2017CQGra..34n4002O},
by marginalizing the likelihood of the data over the seven coordinates characterizing the spacetime coordinates and
orientation of the binary relative to the earth.  
Specifically the likelihood of the data given gaussian noise has the form  (up to normalization)
\begin{equation}
\label{eq:lnL}
\ln {\cal L}(\bm{\lambda} ;\theta )=-\frac{1}{2}\sum\limits_{k}\langle h_{k}(\bm{\lambda} ,\theta )-d_{k} |h_{k}(\bm{\lambda} ,\theta )-d_{k}\rangle _{k}-\langle d_{k}|d_{k}\rangle _{k},
\end{equation}
where $h_{k}$ are the predicted response of the k$^{th}$ detector due to a source with parameters ($\bm{\lambda}$, $\theta$) and
$d_{k}$ are the detector data in each instrument k; $\bm{\lambda}$ denotes the combination of redshifted mass $M_{z}$ and the
remaining parameters needed to uniquely specify the binary's dynamics; $\theta$ represents the
seven extrinsic parameters (4 spacetime coordinates for the coalescence event and 3 Euler angles for the binary's
orientation relative to the Earth); and $\langle a|b\rangle_{k}\equiv
\int_{-\infty}^{\infty}2df\tilde{a}(f)^{*}\tilde{b}(f)/S_{h,k}(|f|)$ is an inner product implied by the k$^{th}$ detector's
noise power spectrum $S_{h,k}(f)$. 
%In all calculations, we adopt the fiducial O1 noise power spectra associated with data near GW150914 \cite{DiscoveryPaper}.
In practice we adopt both  low- and high- frequency cutoffs $f_{\rm max},f_{\rm min}$ so all inner products are modified to
\begin{equation}
\label{eq:overlap}
\langle a|b\rangle_{k}\equiv 2 \int_{|f|>f_{\rm min},|f|<f_{\rm max}}df\frac{[\tilde{a}(f)]^{*}\tilde{b}(f)}{S_{h,k}(|f|)}.
\end{equation}
The joint posterior probability of $\bm{\lambda} ,\theta$ follows from Bayes' theorem:
\begin{equation}
p_{\rm post}(\bm{\lambda} ,\theta)=\frac{ {\cal L}(\bm{\lambda} ,\theta)p(\theta)p(\bm{\lambda})}{\int d\bm{\lambda} d\theta {\cal L}(\bm{\lambda} ,\theta)p(\bm{\lambda})p(\theta)},
\end{equation}
where $p(\theta)$ and $p(\bm{\lambda})$ are priors on the (independent) variables $\theta ,\bm{\lambda}$. For each $\bm{\lambda}$, we evaluate the marginalized likelihood
\begin{equation}
 {\cal L}_{\rm marg}\equiv\int  {\cal L}(\bm{\lambda} ,\theta )p(\theta )d\theta
\end{equation}
via direct Monte Carlo integration, where $p(\theta)$ is uniform in 4-volume and source orientation.  
The integral in time is performed at a low level, by direct quadrature. 
For the remaining dimensions, to evaluate the likelihood in regions of high importance, we use an adaptive Monte Carlo as described in
\cite{gwastro-PE-AlternativeArchitectures}.    As described in Pankow et al, this marginalized likelihood can be evaluated efficiently
because, having generated the dynamics and outgoing radiation in all possible directions once and for all for fixed
$\mathbf{\lambda}$, the likelihood can be evaluated as a function of $\theta$ at very low computational cost.

% From ROM paper
A dimensionless, complex gravitational-wave
strain
%% ~\footnote{\scott{\sout{We use a distance-independent dimensionless strain $R h/M$, where
%% $R$ is the distance from the binary's center-of-mass
%% and $M$ is the total mass. In this paper we choose units so that $c=G=1$.}
%% [Remove this? I think it conflicts with other conventions/definitions 
%% used throughout the paper.]}}
\begin{align} \label{eq:strain}
h(t,\vartheta,\phi;\lambda) =  h_+(t,\vartheta,\phi;\lambda) - 
                                i h_\times (t,\vartheta,\phi;\lambda) \, ,
\end{align}
can be expressed in terms of its two fundamental polarizations $h_+$ and $h_\times$.
Here, $t$ denotes time, $\vartheta$ and $\phi$ are the polar and azimuthal angles
for the direction of gravitational wave propagation away from the source. 
The complex gravitational-wave strain can be written in terms of
spin-weighted spherical harmonics $\Y{-2}_{\ell m} \left(\vartheta, \phi \right)$ as 
\begin{align} \label{eq:strain_mode}
h(t,\vartheta,\phi;\lambda) = 
\sum_{\ell=2}^{\infty} \sum_{m=-\ell}^{\ell} \frac{D_{\rm ref}}{D} h^{\ell m}(t;\lambda) \Y{-2}_{\ell m} \left(\vartheta, \phi \right) \, ,
\end{align}
where the sum includes all available harmonic modes $h^{\ell m}(t;\pmb{\lambda})$ made available by the model;  where
$D_{\rm ref}$ is a fiducial reference distance; and where $D$, the luminosity distance to the  source, is one of the
extrinsic parameters.

Following \citet{gwastro-PE-AlternativeArchitectures}, we substitute expression~\eqref{eq:strain_mode} 
for $h_{\ell m}$ into the expression $h_k(t) =F_{+,k} h_+(t_k) +
  F_{\times,k}h_\times(t_k)$ for the detector response $h_k$, 
where $t_k=t_c - \vec{x}_k \cdot \hat{n}$ is the arrival time at the $k$th detector (at position $\vec{x}_k$)
for a plane wave propagating along $\hat{n}$ \cite{gwastro-PE-AlternativeArchitectures}.
We then substitute these expressions for $h_k$ into the likelihood function~\eqref{eq:loglikelihood}
thereby generating~\cite{gwastro-PE-AlternativeArchitectures}
\begin{widetext}
\begin{align}
\ln {\cal L}(\lambda, \theta) 
&= (D_{\rm ref}/D) \text{Re} \sum_k \sum_{\ell m}(F_k \Y{-2}_{\ell m})^* Q_{k,lm}(\lambda,t_k)\nonumber \\
&   -\frac{(D_{\rm ref}/D)^2}{4}\sum_k \sum_{\ell m \ell' m'}
\left[
{
|F_k|^2 [\Y{-2}_{\ell m}]^*\Y{-2}_{\ell'm'} U_{k,\ell m,\ell' m'}(\lambda)
}
% \right. \nonumber \\ & \left.
 {
+  \text{Re} \left( F_k^2 \Y{-2}_{\ell m} \Y{-2}_{\ell'm'} V_{k,\ell m,\ell'm'} \right)
}
\right]
\label{eq:def:lnL:Decomposed}
\end{align}
\end{widetext}
%% \noindent \scott{[FOR RICHARD: is eq 9/10 missing objects? Looks like $t_c$ an $F_k$ just appeared. On the other hand,
%%     putting them into earlier equations will complicate the discussion and is awkward.]}  
where 
 where $F_k = F_{+,k} - i F_{\times,k}$ are the
complex-valued detector
response functions of the $k$th detector \cite{gwastro-PE-AlternativeArchitectures} and 
the quantities $Q,U,V$ depend on $h$ and the data as
\begin{subequations}
%\label{eq:ComputeRhoViaInnerProductMatrix}
\label{eq:QUV}
\begin{align}
Q_{k,\ell m}(\lambda,t_k) &\equiv \qmstateproduct{h_{\ell m}(\lambda,t_k)}{d}_k \nonumber\\
&= 2 \int_{|f|>f_{\rm low}} \frac{df}{S_{n,k}(|f|)} e^{2\pi i f t_k} \tilde{h}_{\ell m}^*(\lambda;f) \tilde{d}(f)\ , \\
{ U_{k,\ell m,\ell' m'}(\lambda)}& = \qmstateproduct{h_{\ell m}}{h_{\ell'm'}}_k\ , \\
V_{k,\ell m,\ell' m'}(\lambda)& = \qmstateproduct{h_{\ell m}^*}{h_{\ell'm'}}_k  \ .
\end{align}
\end{subequations}
Rewriting sums as matrix operations, the likelihood can be equivalently expressed as
\begin{align}
\ln {\cal L} &= \frac{D_{\rm ref}}{D} \text{Re}[ (F Y)^\dag Q]  \nonumber \\
 & - \frac{D_{\rm ref}^2}{4 D^2} [ (FY)^\dag U FY + (FY)^TV FY]
\label{eq:lnL:MatrixForm}
\end{align}
where $F,D$ are arrays over extrinsic parameters; $Q$ is an array over time and $(l,m)$; and $U,V$ are matrices over
$(l,m)$.

\subsection{Accelerated evaluation via efficient multiplication}
%* GPU code
%* Refactoring code to use the efficient GPU loops

At a fixed set of intrinsic parameters $\lambda$, \ILE{} will repeatedly evaluate the time-marginalized likelihood 
for each candidate Monte Carlo set of extrinsic parameters $\theta$:
\begin{eqnarray}
{\cal L}_{\rm margT} \equiv  \int {\cal L} \frac{dt}{T}
\label{eq:lnL:tmarg}
\end{eqnarray}
where $T$ is a small coincidence window consistent with the candidate event time; here and previously,
$T=0.3\unit{sec}$.   To accelerate the code, after precomputing the inner products $U,V,Q$, we simply shift them to the
graphics card, then carry out all calculations necessary to implement Eqs. (\ref{eq:lnL:MatrixForm}, \ref{eq:lnL:tmarg})
on the GPU.  To enable this implementation with portable code, we use \texttt{cupy}, a drop-in-replacement for equivalent
\texttt{numpy} code used for the CPU-based version of ILE.    For the most costly part of the calculation -- the inner
products necessary to evaluate $Q_{lm}(t)$ -- we use a custom CUDA kernel.  
%
With these changes alone, the likelihood evaluation is roughly $60\times$ faster on equivalent hardware.  
After this update, individual likelihood evaluation costs are not the performance- or cost-limiting feature of RIFT-based source
parameter inference.  Instead, the overhead associated with the adaptive Monte Carlo and with the (CPU-based) inner
product evaluations for $Q,U,V$ dominate our computational cost.

\subsection{Tradeoff between Monte Carlo integration and accuracy}

Each marginalized likelihood evaluation has a relative uncertainty of order  $1/\sqrt{n_{\rm eff}}$, where the number of
effective samples $n_{\rm eff}=\epsilon N_{\rm it}$ increases linearly with the total number $N_{\rm it}$ of Monte Carlo
evaluations performed by \ILE{}.  Therefore, to increase (decrease) our accuracy for each likelihood evaluation by a factor of $A$
will require a factor $A^2$ more (fewer) iterations.   
%
We adopt a fixed threshold on $n_{\rm eff}$.

We do benefit slightly by re-using the adaptive Monte Carlo integrator for each extrinsic point $\lambda$.  Since the
integrator has already identified the likely range of sky locations and distances on the first iteration, each
subsequent evaluation of the marginalized likelihood can converge marginally more quickly.

\section{Validation and Performance}
\label{sec:demo}

We illustrate the code using two synthetic signals: a binary black hole with masses $m_{1,2}=35,30 M_\odot$ and a binary
neutron star with masses $1.4 M_\odot$ and $1.35 M_\odot$.
We generate  synthetic data for a two-detector LIGO configuration, assuming operation at  initial (BBH) or advanced (BNS) design  sensitivity.
Both signals are generated with SEOBNRv4.    We perform a multi-stage iterative RIFT analysis of these signal with the
time-domain SEOBNRv4 (BBH) or TaylorT4 (BNS) approximation, under the simplifying assumption that both objects are point particles with zero
spin.  We repeat this analysis on heterogeneous LVC collaboration computing resources, to assess our variable
performance across architectures.  Notably, we investigate the following GPU options: (a) GTX 1050 Ti at LIGO-CIT,
available in large numbers; (b) V100, available on selected high-performance machines; (c) and
GTX 1050 at LIGO-LHO.    Unless otherwise noted in the text, we will discuss code configurations using CPU-only and GPU (b) using
$4096\unit{Hz}$ sampling and only the  $\ell=2, m=\pm 2$ modes.  We also conservatively report computational cost for
parameters $\lambda$ which are close to or within the posterior and therefore, due to their consistency with the data,
have the best-determined extrinsic parameters and thus converge the most slowly.

\editremark{report on distance to each signal, and in particular $d_{\rm max}$ as that might impact performance too}
chosen to have peak SNR $\simeq 14$ and

\subsection{Massive compact binary black holes}
Previously, each instance of ILE examined one intrinsic point $\lambda$.  The overall cost of this ILE evaluation
involved two parts: a startup cost, a setup cost, and a Monte Carlo integration cost.  The  startup cost $\tau_{start}$
is associated with code setup followed by reading, data conditioning, and Fourier-transforming the
data.  The setup cost $\tau_{setup}$ arises from waveform generation and  the inner products $U,V,Q$.  Finally, the Monte Carlo
cost $\tau_{mc}=N_{ad} \tau_{ad}+N_{it}\tau_{eval}$ increases with the number of Monte Carlo iterations $N_{it}$ with
and the number of times $N_{ad}$ the sampling prior used in adaptive Monte Carlo is regenerated, in proportion to their
cost.   In the standard configuration, the adaptive sampling prior is regenerated every $n_{\rm chunk}$ iterations (i.e,
 $N_{ad} \simeq N_{it}/n_{chunk}$), regenerating a sampling prior in two sky location parameters and distance.  
The choice of one $\lambda$ for each instance of ILE was due to the substantial time  $\tau_{MC}$, which vastly
dominated the overall computational cost.   For example, for a typical analysis of a short signal --  a typical binary
black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$ --  this version has
cost elements as shown in Table \ref{tab:CostBreakdown}, based on the assumption that the MC terminates after $N_{\rm it}\simeq 2\times 10^6$ iterations using
$n_{\rm chunk}\simeq 10^4$.  
The marginalized likelihood described above  converges incredibly rapidly to a small value if the candidate model is
inconsistent with a signal (or the absence thereof) in the data, with only $N_{it} \simeq O(10^4)$  evaluations needed.

\begin{table*}
\begin{tabular}{lrr|ccccc|rr}
Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
$\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
  &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
% ~/parse_report.sh profile_nogpu_pcdev13.log | more
CPU & 16384 & $\pm 2 $ & 20 & &&532 & &  700  \\ 
       & 4096 & $\pm 2 $ &   20 \\ \hline
% ./parse_report.sh 20190130-profile_nogpu_HM_pcdev13.log  | more
%    setup time: PrecomputeLikelihoodTerms, includes waveform generation. 
%   evaluation: FactoredLogLikelihodTimeMarginalized Divide by actual number of calls, since not a block!
%    
CPU & 16384 & $\pm 2,\pm 1 $ & 20 & 1.5 && 680 & &  1000  \\ 
       & 4096 & $ \pm 2, \pm 1 $ &   20 \\ \hline
% LIGO-WA, or a compute node at CIT with a grpahics catd: 
%  - old compute node: one marginalized likelihood in 2 minutes (!), 2x8
%  - CIT jobs, on compute nodes
%        20 points at 16kHz take about 90 min
%        20 points at kHz take anywhere from
%        condor_userlog iteration_0_ile/logs/*.log
GPU (a) & 16384 & $(2,\pm 2) $  & 20  & & && & 270 \\
            & 4096 &$\pm 2 $  &  20 &  & & && 45 \\ \hline
%% % LIGo CIT, pcdev13, 16 kHz : basically indistinguishable from cost at 4kHz
%%       1    0.027    0.027  560.572  560.572 integrate_likelihood_extrinsic_batchmode:8(<module>)
%%        20    0.082    0.004  552.004   27.600 integrate_likelihood_extrinsic_batchmode:690(analyze_event)
%%        20   76.306    3.815  503.342   25.167 mcsampler.py:327(integrate)
%%      2345   57.019    0.024  241.951    0.103 mcsampler.py:198(draw)
%%     25785    0.135    0.000   99.535    0.004 shape_base.py:286(hstack)
%%     26561   99.102    0.004   99.102    0.004 {numpy.concatenate}
%%     21981    0.933    0.000   87.819    0.004 function_base.py:2063(__call__)
%%     21981   34.750    0.002   86.887    0.004 function_base.py:2154(_vectorize_call)
%%      2345    1.391    0.001   86.327    0.037 integrate_likelihood_extrinsic_batchmode:810(likelihood_function)
%%      2345    0.007    0.000   46.648    0.020 __init__.py:630(asnumpy)
%%      2345   46.557    0.020   46.640    0.020 {method 'get' of 'cupy.core.core.ndarray' objects}
%%        20    0.179    0.009   43.145    2.157 factored_likelihood.py:92(PrecomputeLikelihoodTerms)
%%        20    0.174    0.009   37.498    1.875 lalsimutils.py:2582(hlmoft_SEOB_dict)
%%        20    0.099    0.005   37.324    1.866 lalsimutils.py:2743(complex_hoft)
%%        20   37.220    1.861   37.220    1.861 {lalsimulation._lalsimulation.SimInspiralChooseTDWaveform}
%%    444365    0.761    0.000   35.303    0.000 polyint.py:62(__call__)
%%      2345    5.260    0.002   33.956    0.014 factored_likelihood.py:1676(DiscreteFactoredLogLikelihoodViaArrayVectorNoLoop)
%%       407    0.005    0.000   28.757    0.071 odepack.py:28(odeint)
%%       407    1.327    0.003   28.740    0.071 {scipy.integrate._odepack.odeint}
%%    444365    2.644    0.000   26.423    0.000 interpolate.py:657(_evaluate)
%% 117270874   22.569    0.000   22.569    0.000 mcsampler.py:685(uniform_samp)
% LIGO-CIT, pcdev13, 4kHz
%  - 20 iterations, about 450  to 500 sec to eval (i.e., about 25 sec per lnL)
%  -  precompute: can time manually
%   -  Overlap stages: PrecomputeLIkelihood + waveform, about 4
%  - Histogram time: about (12+30) seconds overall (histogram + ODEint)
%  - Actual likelihood evaluation: about 30 sec out of 20 eval, or about 1.5 sec
   %%   1    0.061    0.061  506.265  506.265 integrate_likelihood_extrinsic_batchmode:8(<module>)
   %%     20    0.194    0.010  498.581   24.929 integrate_likelihood_extrinsic_batchmode:690(analyze_event)
   %%     20   80.397    4.020  464.263   23.213 mcsampler.py:327(integrate)
   %%   2140   54.358    0.025  235.480    0.110 mcsampler.py:198(draw)
   %%  23530    0.134    0.000  101.280    0.004 shape_base.py:286(hstack)
   %%  24306  100.861    0.004  100.861    0.004 {numpy.concatenate}
   %%  20136    0.896    0.000   84.698    0.004 function_base.py:2063(__call__)
   %%  20136   33.608    0.002   83.802    0.004 function_base.py:2154(_vectorize_call)
   %%   2140    1.403    0.001   45.791    0.021 integrate_likelihood_extrinsic_batchmode:810(likelihood_function)
   %% 461726    0.790    0.000   36.137    0.000 polyint.py:62(__call__)
   %%   2140    5.496    0.003   30.829    0.014 factored_likelihood.py:1676(DiscreteFactoredLogLikelihoodViaArrayVectorNoLoop)
   %%    407    0.006    0.000   29.662    0.073 odepack.py:28(odeint)
   %%    407    1.360    0.003   29.644    0.073 {scipy.integrate._odepack.odeint}
   %%     20    0.052    0.003   28.785    1.439 factored_likelihood.py:92(PrecomputeLikelihoodTerms)
   %% 461726    2.809    0.000   26.983    0.000 interpolate.py:657(_evaluate)
   %%     20    0.048    0.002   24.454    1.223 lalsimutils.py:2582(hlmoft_SEOB_dict)
   %%     20    0.030    0.002   24.406    1.220 lalsimutils.py:2743(complex_hoft)
   %%     20   24.373    1.219   24.373    1.219 {lalsimulation._lalsimulation.SimInspiralChooseTDWaveform}
% ./parse_report.sh 20190130-profile_gpu_lowres_pcdev13.log_welford | more
% ./parse_report.sh 20190130-profile_gpu_lowres_pcdev13.log_16kHz | more
%   'residual' time: take mcsampler, subtract likleihood, divide by number of calls and number of evals per chunk
GPU (b) & 16384 & $\pm 2$ & 20  & 1.8 & 1 & \editremark{0.85}& \editremark{6} &28 & 15\\
       & 4096 & $\pm 2$  & 20 & $1.2 $ &  1  & \editremark{0.75} & \editremark{6}  & 25\\ \hline
% ~/parse_report.sh profile_gpu_HM_lowres_pcdev13.log | more
%   - extra eval cost: (mcsampler - likelihood)/20/n_eval_likelihood/10^4  
GPU (b) & 16384 & $\pm 2, \pm 1$ & 20 & 1 && 4.2 & \editremark{1}  & 38  \\
       & 4096 & $\pm 2, \pm 1$ & 20 & 1&& 2.5  && 35 & \\ \hline
%
% LHO or LLO worker nodes
%  ./parse_report.sh 20190130-profile_gpu_LHOworker_16kHz.log | more
GPU (c) & 16384 & $\pm 2 $  & 20  & & & 18& &160 &  \\
            & 4096 &$\pm 2 $  &  20 &  & & 22  & & 140 & $\simeq 50$ \\
\end{tabular}
% Compute node at LIGO-WA
\caption{\label{tab:CostBreakdown}\textbf{Profiling performance: Binary black holes}: Evaluation costs for the
  marginalized likelihood on default
  hardware, for a two-mode system $(l,m)=\pm 2$ analyzing $T=8\unit{s}$ of data with a massive binary black hole
  $m_1=35 M_\odot,M_2=30 M_\odot$.  The last column indicates peak GPU utilization.}
\end{table*}

With the new low-cost ${\cal L}_{\rm margT}$ evaluations, however, the startup and setup costs  $\tau_{start},\tau_{setup}$ now can be
comparable or in excess of the total time used to evaluate the marginalized likelihood ${\cal L}_{marg}$.   In fact, the total time per Monte Carlo
evaluation $\tau_{it}=\tau_{manage}+\tau_{eval}$  spent in general-purpose overhead $\tau_{manage}$ will be much larger than the time $\tau_{eval}$ spent
carrying out scientific calculations by evaluating the likelihood.  For these reasons, we reorganize the workflow, so each
instance of ILE loops over
$N_{eval}$ different choices for $\lambda$. 
%
Additionally, particularly for massive binary black holes, we investigate two additional performance improvements.  First and
foremost, we lower the sampling rate, thus lowering the startup cost $\tau_{start}$ and particularly setup cost
$\tau_{setup}$.  Previously and out of an abundance of caution, \ILE{} employed a sampling rate $1/\Delta t = 16384
\unit{Hz}$ for all calculations, but terminated all inner products in Eq. (\ref{eq:overlap}) at roughly $f_{\rm
  max}=2048\unit{Hz}$ or less.   Reducing the sampling rate by a factor $s$ will the  cost of all operations with
timeseries -- they are shorter.  Depending on the relative cost of overhead versus array operations,  $\tau_{setup}$
\emph{and} the cost per evaluation $\tau_{eval}$ decrease modestly because of this effect.  
Also, to insure safe inner products over short data sets in the presence of very narrow spectral lines, we operated on a
significantly-padded data buffer,
using 
``inverse spectrum truncation'' \editremark{citation}.  For binary black holes, the degree of padding was significantly
in excess of the signal duration \editremark{numbers: say 64s for a BBH!}.  
Instead and following LI, we adopt a much shorter inverse spectrum truncation length, a much shorter padding buffer, and
a Tukey window applied to the data to remove discontinuities at the start and end of each segment.   
Reducing the duration of data analyzed by a factor $s'$ will reduce the cost $\tau_{setup}$ by a factor $s'$.  
%
Combining these factors together, the overall computational cost  $T_{ILE}/N_{eval}$  of each marginalized likelihood
evaluation is
\begin{align}
T_{ILE}/N_{\rm eval} &= \frac{\tau_{start}}{N_{\rm eval}} 
 + 
 [ \tau_{setup}/s' + N_{ad} \tau_{ad} + N_{it}\tau_{it}
 ] 
\end{align}
For a typical binary black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$, this
version requires $\tau_{start} \simeq 12\unit{s}$, $\tau_{setup}/ss' \simeq XX$, $\tau_{ad}\simeq XXX$, and $N_{it}\tau_{\it} \simeq XX$
\editremark{numbers}.  
Because of the extremely high cost of adaptive overhead, we only adapt in two parameters, corresponding to sky location; 
and only adapt during the first iteration which returns significant $\ln {\cal L}_{\rm marg}$.  As a result, $N_{ad}$
(the number of adaptive stages \emph{per likelihood evaluation}) scales as $N_{it}/N_{chunk}/N_{\rm eval}$ and does not
increase with the overall number of samples, so the overall evaluation time scales as
\begin{align}
T_{ILE,mod}/N_{\rm eval} &= \frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
 \nonumber \\ &
 + 
 [ \tau_{setup}/s'  + N_{it}\tau_{it}
 ] 
\end{align}
Typically we target $N_{\rm eval} \simeq 20$ so that startup and other one-time costs  are not a significant contributor to the overall
runtime.   In this configuration, the run time per marginalized likelihood is around $T_{ILE}/N_{\rm eval} \simeq 25
\unit{s}$ for a binary black hole.    Of this $25 \unit{s}$, only roughly
$1.5\unit{s}$ corresponds to evaluating the likelihood, implying the cost $\tau_{it} \simeq 1.5/N_{\rm it} \simeq
7.5\times 10^{-7}$.
Likewise, because overhead dominates over array manipulations, an analysis with  $16\unit{kHz}$ timeseries requires only
marginally more time than the corresponding $4\unit{kHz}$ analysis.

For low-latency analyses, we need to assess the overall wallclock time needed to perform an analysis, often with a
restricted number $N_{GPU}$ of available GPU-enabled machines.   To  complete $N_{\rm net}$ likelihood evaluations
with these resources requires
\begin{align}
T_{net,mod} &= \frac{N_{\rm net}}{N_{\rm GPU} N_{\rm eval}} T_{ILE,mod} \nonumber \\
& =
  \frac{N_{net}}{N_{GPU}}\left[
\frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
 \right. \nonumber \\
& \left. + 
 [ \tau_{setup}/ss'  + N_{it}\tau_{it}/s
 ] 
 \right]
\end{align}
Typically we target $N_{\rm eval}  \lesssim 3\times 10^4, N_{\rm GPU} \simeq 100$, or roughly $300\times$ the cost per
individual marginalized likelihood evaluation.  For the binary black hole configuration described above, marginalized
likelihood evaluations will complete in about $100\; \unit{minutes}$ on a cluster of (b), or more realistically 200
minutes on a cluster of (a). . This number can be reduced to
tens of minutes with a modestly larger GPU pool, or by  marginally more conservative convergence thresholds. on $N_{\rm
  eval}$ or $N_{it}$.  
This discussion ignores the latency introduced by the  gaussian process interpolation stage at the end.  We will revisit
accelerated GP interpolation in subsequent work.

The \ILE{} likelihood  generates waveform dynamics and $h_{lm}(t)$ once per evaluation $\lambda$.   Waveform generation can contribute significantly to the
time needed to evaluate  $\ln {\cal L}_{\rm marg}$ for extremely costly
waveform models which require $\tau_{wf} $ a significant fraction of $  \tau_{setup}$.   That said, the tests described
above used relatively costly waveform generation with $h_{lm}(t)$ allocation requiring $\tau_{wf} \simeq 1.2 \unit{s}$
at 4kHz and $1.9\unit{s}$ at 16 kHz.

The \ILE{} likelihood $\ln {\cal L}_{\rm margT}$ involves sums over modes, with the most expensive likelihood and setup operations (involving $Q$) growing linearly with
the number of modes used.  This increased cost is most apparent in Table \ref{tab:CostBreakdown} when comparing the
columns corresponding  $\tau_{it,like}$ for GPU (b) between the $\pm 2$ and $(\pm 2,\pm 1)$ rows.  The cost of
evaluating $\ln {\cal L}_{\rm margT}$   with 4 modes is roughly twice that of runs with two modes, as expected.

In a heterogeneous computing environment, code performance varies substantially with available resources, as seen by
contrasting corresponding results for (a,b,c) in Table \ref{tab:CostBreakdown}.      For
the slowest hardware,  the overall runtime will be only a factor few smaller than the corresponding CPU-only runtime.

\subsection{Binary neutron stars}
\begin{table*}
\begin{tabular}{lrr|ccccc|rr}
Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
$\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
  &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
% ./parse_report.sh 20190130-profile_bns_nogpu_16kHz_pcdev13.log | more
%   - resampling time: take (mcsampler - likelihood), divide by 20=Neval, divide by number of adaptive steps (100)
%   (11261 - 10974)/20/100. = 0.14 sec for 16 kHz
CPU & 16384 & $\pm 2,\pm 1 $ & 35 & 26 & 0.14  & 680  & & 590    \\ 
       & 4096 & $\pm 2, \pm 1$ & 35 & &  &   &&  & \\ \hline
GPU (a) & 16384 & $\pm 2, \pm 1$   & 35 &  &&&&   \\
       & 4096 & $ \pm 2, \pm 1 $        &  35  & &&&& 360 \\ \hline
%   - resampling time: take (mcsampler - likelihood), divide by 20=Neval, divide by number of adaptive steps (100)
%   (226.489 - 17)/20/100  : 0.1 sec for 4kHz
%   (165 - 29.  )/20/100   : looks odd for 16 kHz
GPU (b) & 16384 & $\pm 2, \pm 1$   & 35 & 26 & 0.07 &  2.4 & &35 &  \\
       & 4096 & $ \pm 2, \pm 1 $        &  35  & 11.5 & 0.1 &  2.4  & &24 & \\ \hline
\end{tabular}
\caption{\label{tab:CostBreakdown:BNS}\textbf{Profiling performance: Binary neutron stars}: Evaluation costs for the
  marginalized likelihood on default
  hardware,  analyzing $T=8\unit{s}$ of data with a binary neutron stars
  $m_1=1.4 M_\odot,M_2=1.35 M_\odot$, with convergence threshold $n_{\rm eff} > 50$.
}
\end{table*}

Table \ref{tab:CostBreakdown:BNS} shows the corresponding performance breakdown for our fiducial binary neutron star,
which has signal SNR=32.  In this analysis, we have adopted the TaylorT4 model with $\ell=2$ and all $m=\pm 2,\pm 1$ modes as our
fiducial analysis template.   Despite using more modes and longer signals, and despite an overall likelihood evaluation
cost $\tau_{it}$ consistent BBH analyses,  likelihood evaluations for BNS are often
\emph{as fast or faster} than the corresponding likelihood evaluations for BBH with otherwise comparable settings.

Low-mass compact binaries like binary neutron stars have much longer inspiral times from the same starting frequency of
$20\unit{Hz}$: roughly $160 \unit{s}$.   Nominally, the manipulation of a corresponding power-of-two duration data ($256\unit{s}$) involves
$32\times$ more costly Fourier transforms and time integrations than the BBH analysis described above.  
%  a = np.zeros(16384*256,dtype=np.complex64); a.nbytes/1024/1024
[An array of $256\unit{s}$ of data at $16\unit{kHz}$ corresponds to  32 Mb.]  Comparing $\tau_{setup}$ appearing in Tables
 \ref{tab:CostBreakdown} and  \ref{tab:CostBreakdown:BNS}, we indeed find our BNS analysis requires  a factor of order   $32\times$
 longer to set up all necessary inner products.  Unlike the BBH analysis, the inner product evaluation costs in
 $\tau_{setup}$ now dominates the
 overall evaluation time $T_{ILE}/N_{\rm eval}$.

Improved performance arises instead from re-using our adaptive integrator, which (after the first marginalized
likelihood evaluation) can exploit tight localization afforded by the many BNS cycles available in band and the high
amplitude of our fiducial BNS signal.  In other
words, the Monte Carlo integral converges to our target accuracy in of order $10\times $ fewer steps than for BBH.  
Because this improved performance relies on the sampler, not on our choice of hardware, we achieve a similar substantial
improvement for all choices of GPU.

\editremark{test on other hardware -- also problem, a lot of the speed improvement occurs because the initial grid is
  crap/too wide}

For the closed-form post-Newtonian approximation used in the study above, waveform costs are a modest contribution to
overall runtime, contributing only 4.1 sec to $\tau_{setup}$ at 16 kHz.
For other approximations not available in closed form, the waveform generation cost for binary neutron star inspirals
can be much more substantial, particularly as $f_{\rm min}$ decreases below the 20\unit{Hz} used in this study.

\section{ Results (optional?)}

* Add Jake as author.  Rerun all O2 BBHs with SEOBNRv3?

\section{Conclusions}
We have demonstrated that the marginalized likelihood $\ln {\cal L}_{\rm marg}(\lambda)$ appearing in the RIFT/rapidPE parameter
inference calculation can be evaluated at fixed $\lambda$ in tens of seconds on average for
both binary black holes and binary neutron stars.  This performance improvement could enable very low latency source
parameter inference for compact binaries, which can be of use for targeting multimessenger followup observations via sky
localization and  precise source characterization.     This prospect is particularly interesting because RIFT can
often achieve this performance using computationally costly models for binary merger, as the waveform generation cost
does not usually limit code performance.

The overall code performance and thus latency can be further decreased substantially, notably by converting the Monte
Carlo random number generation and inner products to GPU-based operations.  In such a configuration, the marginalized
likelihood would carry out almost all calculations (except waveform generation) on a GPU, with minimal communication off
the board.  This configuration should further reduce the average time needed to compute $\ln {\cal L}_{\rm
  marg}(\lambda)$ for both binary black holes (which are Monte Carlo limited) and binary neutron stars (which are
inner-product limited).   We anticipate a further factor of roughly 10 reduction in overall evaluation time can be
achieved soon.  At that level of performance, a single 8-GPU machine with contemporary hardware could perform parameter inference less than 10
minutes.    Since binary compact objects intersect our past light cone only once every roughly 15 minutes, accounting
for all past history, such a configuration would be able to address low-latency parameter inference for the duration of
2nd-generation ground-based observing.
% based on Neval ~ 10^4

The use of \texttt{cupy} enables our code to be highly portable across architectures and heterogeneous GPU environments,
while transitioning smoothly between GPU and CPU mode.     The techniques we used here will be transported to other
Bayesian inference modeling codes used to interpret GW observations.
% Will swap version of code

\begin{acknowledgements}
RO'S is supported by NSF PHY-1707965 and PHY-1607520.
DW thanks the RIT COS and CCRG for support, and  NSF-1707965.
Y-LLF was supported in part by BNL LDRD projects No.\,17-029 and No.\,19-002, 
and New York State Urban Development Corporation, d/b/a Empire State Development, under contract No.\,AA289.
We thank the Computational Science Initiative at BNL for hosting GPU Hackathon 2018, during which a part of this work was conducted. We also thank Jeff Layton at NVIDIA for bringing CuPy to our attention.
\end{acknowledgements}

%\bibstyle{unsrt}
\bibliography{paperexport,LIGO-publications}
\end{document}